Dirk Sliwka

### 2025-03-21 about.qmd
Update bib

---
#title: "Dirk Sliwka"
---
<br>
<span style="font-size: 1.5em;">**Dirk Sliwka**</span><br>
[University of Cologne](https://wiso.uni-koeln.de/en/)

::: {.grid}

::: {.g-col-8}
I am a professor of management in the Faculty of Management, Economics, and Social Sciences at the [University of Cologne](https://wiso.uni-koeln.de/en/) in the department [Corporate Development](https://cd.wiso.uni-koeln.de/en/). I am also senior research fellow at the [Max Planck Institute for Research on Collective Goods](https://www.coll.mpg.de/), a research fellow at [IZA](https://www.iza.org/person/790/dirk-sliwka) and [CESifo](https://www.cesifo.org/en), and a member of [C-SEB](https://c-seb.de/) and [ECONtribute](https://econtribute.de/).<br>
<br>
My research focuses on designing effective Human Resource Management practices within firms and evaluating their impact on behavior, performance, and well-being. I am particularly interested in the design of incentives and performance management processes as well as the role of social norms in organizations. <br>
<br>
My main research fields of interest are:<br>
<span style="display: block; margin-bottom: 6pt;"></span>
<ul>
    <li>Organizational/Personnel Economics</li>
    <li>Human Resource Management</li>
    <li>Field Experiments in Firms</li>
    <li>Behavioral Economics</li>
    <li>Management Accounting</li>
</ul>
:::

::: {.g-col-4}
![](assets/img/SliwkaDirk.jpg)
<br>
<div style="font-size: 1em; margin-top: 6px;">
    University of Cologne
    Universitätsstraße 22a
    D-50937 Köln
    Germany
</div>

:::

:::

>>>>

---
title: "Dirk Sliwka"
about:
  template: solana
  image: assets/img/SliwkaDirk.jpg
---

I am a professor of management in the Faculty of Management, Economics, and Social Sciences at the [University of Cologne](https://wiso.uni-koeln.de/en/) in the department [Corporate Development](https://cd.wiso.uni-koeln.de/en/). I am also senior research fellow at the [Max Planck Institute for Research on Collective Goods](https://www.coll.mpg.de/), a research fellow at [IZA](https://www.iza.org/person/790/dirk-sliwka) and [CESifo](https://www.cesifo.org/en), and a member of [C-SEB](https://c-seb.de/) and [ECONtribute](https://econtribute.de/).<br>
<br>
My research focuses on designing effective Human Resource Management practices within firms and evaluating their impact on behavior, performance, and well-being. I am particularly interested in the design of incentives and performance management processes as well as the role of social norms in organizations. <br>
<br>
My main research fields of interest are:<br>
<span style="display: block; margin-bottom: 6pt;"></span>
<ul>
    <li>Organizational/Personnel Economics</li>
    <li>Human Resource Management</li>
    <li>Field Experiments in Firms</li>
    <li>Behavioral Economics</li>
    <li>Management Accounting</li>
</ul>

### 2025-03-21 papers.qmd
Update bib

@@

@@

<a href="assets/pdf/OSW_Final_Paper_13Feb2024.pdf" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  PDF  </a>&nbsp;&nbsp;<button onclick="toggleAbstract('abstract-2024-Multi-rater-Performance-Evaluations-and-Incentives')" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; background: none; display: inline-block;">Abstract</button>
<div id="abstract-2024-Multi-rater-Performance-Evaluations-and-Incentives" style="display:none; border: 1px solid #ccc; padding: 10px; margin-top: 10px;">We compare evaluations of employee performance by individuals and groups of supervisors, analyzing a formal model and running a laboratory experiment. The model predicts that multi-rater evaluations are more precise than single-rater evaluations if groups rationally aggregate their signals about employee performance. Our controlled laboratory experiment confirms this prediction and finds evidence that this can indeed be attributed to accurate information processing in the group. Moreover, when employee compensation depends on evaluations, multi-rater evaluations tend to be associated with higher performance.</div>
<a href="assets/pdf/SocPrefEval2023.pdf" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  PDF  </a>&nbsp;&nbsp;<a href="https://doi.org/10.1287/mnsc.2022.02267" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  WEB  </a>&nbsp;&nbsp;<button onclick="toggleAbstract('abstract-2024-Social-Preferences-and-the-Informativeness-of-Subjective-Performance-Evaluations')" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; background: none; display: inline-block;">Abstract</button>
<div id="abstract-2024-Social-Preferences-and-the-Informativeness-of-Subjective-Performance-Evaluations" style="display:none; border: 1px solid #ccc; padding: 10px; margin-top: 10px;">Abstract We study biases and the informativeness of subjective performance evaluations in an MTurk experiment, testing the implications of a standard formal framework of rational subjective evaluations. In the experiment, subjects in the role of workers perform a real effort task. Subjects in the role of supervisors observe samples of the workers’ output and assess their performance. We conduct 6 experimental treatments varying (i) whether workers’ pay depends on the performance evaluation, (ii) whether supervisors are paid for the accuracy of their evaluations, and (iii) the precision of the information available to supervisors. Moreover, we use the exogenous assignment of supervisors to workers to investigate the association between supervisors' social preferences and their rating quality. In line with the model of optimal evaluations, we find that ratings are more lenient and less informative when they determine bonus payments. Rewards for accuracy reduce leniency and can enhance informativeness. When supervisors have access to more detailed performance information, their ratings vary more with the performance signal and become more informative. Contrary to expectations, we do not find that more prosocial supervisors are systematically more lenient when their ratings affect worker's payoffs. Instead, they are more diligent in their rating behavior, resulting in more accurate and informative performance evaluations.</div>
<a href="assets/pdf/Assignment2023.pdf" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  PDF  </a>&nbsp;&nbsp;<a href="https://doi.org/10.1287/mnsc.2022.03362" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  WEB  </a>&nbsp;&nbsp;<button onclick="toggleAbstract('abstract-2024-The-Algorithmic-Assignment-of-Incentive-Schemes')" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; background: none; display: inline-block;">Abstract</button>
<div id="abstract-2024-The-Algorithmic-Assignment-of-Incentive-Schemes" style="display:none; border: 1px solid #ccc; padding: 10px; margin-top: 10px;">The assignment of individuals with different observable characteristics to different treatments is a central question in designing optimal policies. We study this question in the context of increasing workers' performance via targeted incentives, using machine learning algorithms with worker demographics, personality traits, and preferences as input. Running two large-scale experiments we show that (i) performance can be predicted by accurately measured worker characteristics, (ii) a machine learning algorithm can detect heterogeneity in responses to different schemes, (iii) a targeted assignment of schemes to individuals increases performance significantly above the level of the single best scheme, and (iv) algorithmic assignment is more effective for workers who have a high likelihood to repeatedly interact with the employer, or who provide more consistent survey answers.</div>

>>>>

## 2025
**Managing Skills in Organizations-Evidence from a Field Experiment**<br>
<a href="https://leonhardgrabe.github.io/">Leonhard Grabe</a>, <a href="#">Dirk Sliwka</a><br>
*IZA Discussion Paper 17727*<br>
<button onclick="toggleAbstract('abstract-2025-Managing-Skills-in-Organizations-Evidence-from-a-Field-Experiment')" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; background: none; display: inline-block;">Abstract</button>
<div id="abstract-2025-Managing-Skills-in-Organizations-Evidence-from-a-Field-Experiment" style="display:none; border: 1px solid #ccc; padding: 10px; margin-top: 10px;">We study the value of skill management in organizations. In a natural field experiment with 2,582 service technicians, we vary managers’ ability to monitor and manage employee skills. We find that removing managers’ access to hard information on employee skills reduced training intensity, work performance, and job satisfaction. Combining detailed personnel records and survey data, we show that the intervention lowered employee efforts to identify training needs and managerial attention to employee development. In particular, high-skill employees received less training to broaden their skill set and, in turn, performance losses are driven by higher completion times for complex work assignments.</div>
<a href="assets/pdf/OSW_Final_Paper_13Feb2024.pdf" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  PDF  </a>&nbsp;&nbsp;<button onclick="toggleAbstract('abstract-2025-Multi-rater-Performance-Evaluations-and-Incentives')" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; background: none; display: inline-block;">Abstract</button>
<div id="abstract-2025-Multi-rater-Performance-Evaluations-and-Incentives" style="display:none; border: 1px solid #ccc; padding: 10px; margin-top: 10px;">We compare evaluations of employee performance by individuals and groups of supervisors, analyzing a formal model and running a laboratory experiment. The model predicts that multi-rater evaluations are more precise than single-rater evaluations if groups rationally aggregate their signals about employee performance. Our controlled laboratory experiment confirms this prediction and finds evidence that this can indeed be attributed to accurate information processing in the group. Moreover, when employee compensation depends on evaluations, multi-rater evaluations tend to be associated with higher performance.</div>
<a href="assets/pdf/SocPrefEval2023.pdf" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  PDF  </a>&nbsp;&nbsp;<a href="https://doi.org/10.1287/mnsc.2022.02267" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  WEB  </a>&nbsp;&nbsp;<button onclick="toggleAbstract('abstract-2025-Social-Preferences-and-the-Informativeness-of-Subjective-Performance-Evaluations')" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; background: none; display: inline-block;">Abstract</button>
<div id="abstract-2025-Social-Preferences-and-the-Informativeness-of-Subjective-Performance-Evaluations" style="display:none; border: 1px solid #ccc; padding: 10px; margin-top: 10px;">Abstract We study biases and the informativeness of subjective performance evaluations in an MTurk experiment, testing the implications of a standard formal framework of rational subjective evaluations. In the experiment, subjects in the role of workers perform a real effort task. Subjects in the role of supervisors observe samples of the workers’ output and assess their performance. We conduct 6 experimental treatments varying (i) whether workers’ pay depends on the performance evaluation, (ii) whether supervisors are paid for the accuracy of their evaluations, and (iii) the precision of the information available to supervisors. Moreover, we use the exogenous assignment of supervisors to workers to investigate the association between supervisors' social preferences and their rating quality. In line with the model of optimal evaluations, we find that ratings are more lenient and less informative when they determine bonus payments. Rewards for accuracy reduce leniency and can enhance informativeness. When supervisors have access to more detailed performance information, their ratings vary more with the performance signal and become more informative. Contrary to expectations, we do not find that more prosocial supervisors are systematically more lenient when their ratings affect worker's payoffs. Instead, they are more diligent in their rating behavior, resulting in more accurate and informative performance evaluations.</div>
<a href="assets/pdf/Assignment2023.pdf" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  PDF  </a>&nbsp;&nbsp;<a href="https://doi.org/10.1287/mnsc.2022.03362" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  WEB  </a>&nbsp;&nbsp;<button onclick="toggleAbstract('abstract-2025-The-Algorithmic-Assignment-of-Incentive-Schemes')" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; background: none; display: inline-block;">Abstract</button>
<div id="abstract-2025-The-Algorithmic-Assignment-of-Incentive-Schemes" style="display:none; border: 1px solid #ccc; padding: 10px; margin-top: 10px;">The assignment of individuals with different observable characteristics to different treatments is a central question in designing optimal policies. We study this question in the context of increasing workers' performance via targeted incentives, using machine learning algorithms with worker demographics, personality traits, and preferences as input. Running two large-scale experiments we show that (i) performance can be predicted by accurately measured worker characteristics, (ii) a machine learning algorithm can detect heterogeneity in responses to different schemes, (iii) a targeted assignment of schemes to individuals increases performance significantly above the level of the single best scheme, and (iv) algorithmic assignment is more effective for workers who have a high likelihood to repeatedly interact with the employer, or who provide more consistent survey answers.</div>

@@

**When bonuses backfire: evidence from the workplace**<br>
<a href="https://sites.google.com/view/jakobalfitian">Jakob Alfitian</a>, <a href="#">Dirk Sliwka</a>, <a href="https://sites.google.com/site/vogelsangtimo/">Timo Vogelsang</a><br>
<a href="assets/pdf/Backfire23.pdf" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  PDF  </a>&nbsp;&nbsp;<a href="https://doi.org/10.1287/mnsc.2022.00484" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; display: inline-block;">  WEB  </a>&nbsp;&nbsp;<button onclick="toggleAbstract('abstract-2024-When-bonuses-backfire:-evidence-from-the-workplace')" style="border: 1px solid #007a04; color: #007a04; padding: 1px 8px; text-decoration: none; font-size: 0.7em; background: none; display: inline-block;">Abstract</button>
<div id="abstract-2024-When-bonuses-backfire:-evidence-from-the-workplace" style="display:none; border: 1px solid #ccc; padding: 10px; margin-top: 10px;">Monetary incentives are widely used to align employee actions with employer objectives. We conducted a field experiment in a retail chain to evaluate whether an attendance bonus could reduce employee absenteeism. Apprentices in 232 stores were randomly assigned to a control group or one of two treatment groups in which a monetary or time-off attendance bonus was introduced for one year. We find that neither variant of the attendance bonus led to a systematic reduction in absenteeism. On the contrary, the monetary attendance bonus increased absenteeism substantially by around 50% on average, which corresponds to more than five additional days absent per employee and year. This effect was driven by the most recently hired apprentices. Survey results reveal that the monetary attendance bonus shifted the perception of absenteeism as acceptable behavior. The backfiring effect persisted beyond the end of the experiment, indicating a lasting erosion of social norms.</div>

@@
